---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo=FALSE, warning=FALSE, message=FALSE}
message("The following plots will show properly only if the FEV1 validation copy of epicR - where smoking status change is disabled - is executed. For a permanent copy of the older version of these results, please refer to the report dated 2018-02-15")
library (epicR)
library(ggplot2)


fev1_projection_amin <- function(fev1_0, time, sex, smoking, age, weight, height){
  
  x<-c(0:time)
  
  if (sex==0){
    gender<-1
  } else if (sex==1){
    gender<-2
  }

  v_t<-0.000753  #random effect variance for rate of decline
  v_b<- 0.09711 #random effect variance for baseline
  v_cov<- 0.00062 #covariance, based on COLD and Zafar
  
  fev1_avg <- c()
  fev1_up <- c()
  fev1_low <- c()
  fev1_annual_rate <- c()
  
     fev1_0_ZafarCMAJ_by_sex <- cbind(male = c(intercept = 1.4275 + 0.4825, baseline_age = -0.00508, baseline_weight_kg = -0.00049, height = -1.8725, height_sq = 1.9513, current_smoker = -0.09221, age_height_sq = -0.00832, followup_time = 0),
                                    female = c(intercept = 1.4275,  baseline_age = -0.00508, baseline_weight_kg = -0.00049, height = -1.8725, height_sq =  1.9513, current_smoker = -0.09221, age_height_sq = -0.00832, followup_time = 0))
     
     

     fev1_betas_by_sex <- cbind(male = c(intercept = -0.1543 - 0.00762, baseline_age = 0.002344, baseline_weight_kg = 0.000126, height =0, height_sq = 0, current_smoker = -0.03074, age_height_sq = 0, followup_time =-0.00146),
                            female = c(intercept = -0.1543 , baseline_age = 0.002344, baseline_weight_kg =0.000126, height =0, height_sq =0, current_smoker =  -0.03074, age_height_sq = 0, followup_time =-0.00146)) #debug
     


   dfev1_re_rho <- 0.09956332 # From G Matrix: 0.00087/(sqrt(0.1006)*sqrt(0.000759))
   dfev1_sigmas <- t(as.matrix(c(sigma1 = sqrt(0.1006), sigma2 = sqrt(0.000759))))
   fev1_0_inc_sd_by_sex <- c(male = 0.54, female = 0.36)
     

  
       beta_0 <- fev1_0-
    ( fev1_0_ZafarCMAJ_by_sex[2, gender]*age + 
        fev1_0_ZafarCMAJ_by_sex[3, gender]* weight + 
        fev1_0_ZafarCMAJ_by_sex[4, gender]* height +
        fev1_0_ZafarCMAJ_by_sex[5, gender]* height* height +
        fev1_0_ZafarCMAJ_by_sex[6, gender]* smoking +
        fev1_0_ZafarCMAJ_by_sex[7, gender]*age* height* height)


       beta_0_prime <-   fev1_betas_by_sex[1, gender]+ dfev1_sigmas[2]/ dfev1_sigmas[1]* dfev1_re_rho*(beta_0 - fev1_0_ZafarCMAJ_by_sex[1, gender])
       
       fev1_slope <- beta_0_prime +
          fev1_betas_by_sex[2, gender]*age +
          fev1_betas_by_sex[3, gender]* weight +
          fev1_betas_by_sex[4, gender]* height +
          fev1_betas_by_sex[5, gender]* height* height +
          fev1_betas_by_sex[6, gender]* smoking +
          fev1_betas_by_sex[7, gender]*age* height* height;

        fev1_slope_t= fev1_betas_by_sex[8, gender];

         fev1_avg <- fev1_0 + time*fev1_slope + time*time*fev1_slope_t
         SE_fev1_slope <- sqrt((1- dfev1_re_rho* dfev1_re_rho)* dfev1_sigmas[2]* dfev1_sigmas[2])
         SE_fev1 <- time * SE_fev1_slope
         
        fev1_up <- fev1_avg+1.96*(SE_fev1)
        fev1_low <- fev1_avg-1.96*(SE_fev1)
   

  df <-data.frame(fev1_slope, fev1_low, fev1_up,  fev1_avg)
  names(df) <- c("FEV1_rate", "FEV1_lower", "FEV1_upper", "final_FEV1" )
  
  return(df)
  
}

init_session()
run(50000)
events <- as.data.frame (Cget_all_events_matrix())
terminate_session()


### annual by LHS for male smokers ;

IDs <- subset (events, (local_time==0) & LHS_eligible==1 & smoking_status==1 & sex ==0 & gold>0)
annualEvents <- subset (events, id %in% IDs$id & event==1)
annualEvents <- subset (annualEvents, gold!=0 & followup_after_COPD > 2) #excluding non-COPD cases

PI2 <- matrix (NA, nrow = nrow(annualEvents), ncol = 3)
PI2 <- as.data.frame(PI2)

for (i in 1:nrow(annualEvents)) {
  tmp <- fev1_projection_amin(annualEvents[i, "FEV1_baseline"], annualEvents[i, "local_time"], annualEvents[i,"sex"], annualEvents[i,"smoking_status"], annualEvents[i,"age_baseline (COPD)"], annualEvents[i,"weight_at_COPD"], annualEvents[i,"height"])
  
  PI2 [i,"LHS_expected_FEV1_low"] <- tmp[["FEV1_lower"]]
  
  PI2 [i,"LHS_FEV1_rate"] <- tmp[["FEV1_rate"]]
  
  PI2 [i,"LHS_expected_FEV1_up"] <- tmp[["FEV1_upper"]]
  
  PI2 [i,"LHS_expected_FEV1"]<- tmp[["final_FEV1"]]
  
}

results <- data.frame(annualEvents$id, annualEvents$followup_after_COPD , annualEvents$FEV1_baseline,annualEvents$FEV1_slope, annualEvents$FEV1,  PI2$LHS_FEV1_rate, PI2$LHS_expected_FEV1_low, PI2$LHS_expected_FEV1_up, PI2$LHS_expected_FEV1)
colnames(results) = c("id", "year", "fev1_baseline", "epicR_FEV1_rate", "epicR_predicted_FEV1", "LHS_FEV1_rate", "LHS_expected_FEV1_low", "LHS_expected_FEV1_up", "LHS_expected_FEV1")

for (i in 1:length(results$epicR_predicted_FEV1)){
  if (results$epicR_predicted_FEV1[i] < results$LHS_expected_FEV1_low[i]) {results$WithinPI[i] <- "EPIC Underestimates"}
  else if (results$epicR_predicted_FEV1[i] > results$LHS_expected_FEV1_up[i]) {results$WithinPI[i] <- "EPIC Overestimates"}
  else {results$WithinPI[i] <- "WithinPI"}
}
print (results$WithinPI[i])

results <- subset(results, (epicR_predicted_FEV1 > 0.5 & LHS_expected_FEV1 > 0.5)) #removing negative values

ggplot(results, aes(x = LHS_expected_FEV1   , y = epicR_predicted_FEV1)) +
  geom_point (size = 1, aes(colour = factor(WithinPI))) + labs(title = "Lung Function", caption = "Continuous smoker males over 40 years with COPD. ", y = "EPIC-Simulated FEV1", x = "LHS Model FEV1") + geom_ribbon(aes(ymin=(LHS_expected_FEV1_low), ymax= (LHS_expected_FEV1_up)), linetype=2, alpha=0.3) + xlim(1,4) + ylim(1,4)

ggsave(paste0("FEV1_validation_male_smoker",".tiff"), plot = last_plot(), device = "tiff", dpi = 300)

message ("Zafari model with all covariates. Continuous smoker males with COPD over 40 years of age")
results$WithinPI <- as.factor(results$WithinPI)
summary(results)
LHS <- results$LHS_expected_FEV1
epic <- results$epicR_predicted_FEV1
dif2 <- var(epic-LHS)
r2 <- 1 - (dif2/var(LHS))
r2

summary(results)

