---
title: "EPIC Background"
output: 
  html_document:
    toc: true
    toc_float: true  
    toc_depth: 4
    css: []

vignette: >
  %\VignetteIndexEntry{BackgroundEPIC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(rmarkdown.html_vignette.check_title = FALSE)
```


## 1 Introduction   

The Evaluation Platform In COPD (EPIC) was a nationally funded research project with the aim of
creating an open-source, publicly available, population-based ‘Whole Disease’ COPD model for
epidemiological projections and policy analyses in the Canadian context.

## 2 The Study 
The **epicR** package and application are based on the paper [Development and Validation of theEvaluation Platform in COPD (EPIC):A Population-Based Outcomes Modelof COPD for Canada](https://journals.sagepub.com/doi/pdf/10.1177/0272989X18824098), published in 2019.

### 2.1 Background

The purpose of this study is to model how COPD (Chronic Obstructive Pulmonary Disease) affects a whole population. In social sciences and population health, we often use a **dynamic microsimulation** model to look at how a disease might affect a population over a period of time.

### What is a Dynamic Microsimulation? {-}

A dynamic microsimulation is a computer model which makes predictions about how a population will be affected by say, a disease, over a set period of time.

In the case of **EPIC**, we wanted to look at the affects of COPD on the Canadian population. To start with, the computer model uses data from the [2001 Canadian Community Health Survey 1.1](http://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&Id=3359) to create simulated, or "artificial" people. The survey included information from 105 908 Canadians. The model uses this to create ~ 22.5 million community-dwelling Canadians.

As with any real population, the simulated population is constantly changing. We define a change in a simulated individual as an **event**.


```{r, eval=TRUE, echo=FALSE, warning=FALSE}
library(htmltools)
includeCSS("html/section-2/event-table.css")
rawHTML <- paste(readLines("html/section-2/event-table.html"), collapse="\n")
HTML(rawHTML)
```


We combined 12 different studies to look at 7 different aspects of COPD:

1) [Demographic and Risk Factor Module](#demographic)
2) [COPD Occurence Module](#occurence)
3) [Lung Function Module](#lungfunction)
4) [Exacerbation Module](#exacerbation)
5) [Mortality Module](#mortality)
6) [Payoff Module](#payoff)
7) [Smoking Module](#smoking)

### Demographic and Risk Factor Module {#demographic}

The Demographic and Risk Factor Module is based on the results from [POHEM](https://pophealthmetrics.biomedcentral.com/track/pdf/10.1186/s12963-015-0057-x), which is the dynamic microsimulation model we described previously. This is basically the background data for the simulation.


```{r, eval=TRUE, echo=FALSE, warning=FALSE}
library(htmltools)
includeCSS("html/section-2/dropdown-table.css")
tags$script(src = "html/section-2/toggle-buttons.js")
rawHTML <- paste(readLines("html/section-2/demographic-table.html"), collapse="\n")
HTML(rawHTML)
```

### COPD Occurence Module {#occurence}

We used data from the Canadian Cohort of Obstructive Lung Disease (COLD) to assign a binary COPD status to individuals upon their creation.

```{r, eval=TRUE, echo=FALSE, warning=FALSE}
library(htmltools)
includeCSS("html/section-2/dropdown-table.css")
rawHTML <- paste(readLines("html/section-2/occurrence-table.html"), collapse="\n")
HTML(rawHTML)

```

### Lung Function Module {#lungfunction}

Once the COPD designation is defined for an individual, an individual-specific initial FEV₁  value and an individual-specific annual rate of FEV₁  decline are assigned. The 3 components of this module are the initial FEV₁  value for preexisting (prevalent) COPD cases, initial FEV₁  values for incident COPD cases, and the slope of decline in FEV₁ over time.

```{r, eval=TRUE, echo=FALSE, warning = FALSE}
library(htmltools)
includeCSS("html/section-2/dropdown-table.css")
rawHTML <- paste(readLines("html/section-2/lungfunction-table.html"), collapse="\n")
HTML(rawHTML)
```

### Exacerbation Module {#exacerbation}

In COPD, patients can sometimes experience what are called exacerbations.

**exacerbation**: acute worsening of COPD symptoms; acute means sudden onset, short in duration; COPD symptoms include shortness of breath, wheezing, coughing up mucus, etc.

#### Exacerbation Severity

```{r, eval=TRUE, echo=FALSE, warning=FALSE}
library(htmltools)
rawHTML <- paste(readLines("html/section-2/exacerbation-severity.html"), collapse="\n")
HTML(rawHTML)


```

```{r, eval=TRUE, echo=FALSE, warning=FALSE}
library(htmltools)
includeCSS("html/section-2/dropdown-table.css")
rawHTML <- paste(readLines("html/section-2/exacerbation-table.html"), collapse="\n")
HTML(rawHTML)
```

 **hazard**: instantaneous exacerbation rate

### Mortality Module {#mortality}

We categorized death into two categories: death from COPD, and death not from COPD. More formally:

**COPD-related mortality**: Death due to a severe or very severe COPD exacerbation.
**Background mortality**: Death from all other causes (excluding COPD).

```{r, eval=TRUE, echo=FALSE, warning=FALSE}
library(htmltools)
includeCSS("html/section-2/dropdown-table.css")
rawHTML <- paste(readLines("html/section-2/mortality-table.html"), collapse="\n")
HTML(rawHTML)
```

### Payoff Module {#payoff}

Like any disease, COPD costs money to treat, and is a cost on the healthcare system. Additionally, having COPD can reduce quality of life, which depends on the severity of the diagnosis and number of exacerbations. We define cost and utility, below, and divided each into 2 categories:

```{r, eval=TRUE, echo=FALSE, warning=FALSE}
library(htmltools)
includeCSS("html/section-2/dropdown-table.css")
rawHTML <- paste(readLines("html/section-2/payoff-breakdown.html"), collapse="\n")
HTML(rawHTML)
```

#### Module Summary

```{r, eval=TRUE, echo=FALSE, warning=FALSE}
library(htmltools)
includeCSS("html/section-2/dropdown-table.css")
rawHTML <- paste(readLines("html/section-2/payoff-table.html"), collapse="\n")
HTML(rawHTML)
```

### Smoking Module {#smoking}

This module assesses the impact of smoking on the development and progression of COPD. Smoking status (e.g., smoker, former smoker, non-smoker) is incorporated into the model to account for its effect on the individual’s health trajectory.

```{r, eval=TRUE, echo=FALSE, warning=FALSE}
library(htmltools)
includeCSS("html/section-2/dropdown-table.css")
rawHTML <- paste(readLines("html/section-2/smoking-table.html"), collapse="\n")
HTML(rawHTML)
```

