---
title: "Payoffs"
output: html_document
vignette: >
  %\VignetteIndexEntry{Payoffs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

This document outlines the steps taken to determine input values for the payoff module. 

## Step 1: Estimating annual COPD medication costs

Annual COPD medication costs were estimated using two components:

1. **Inhaler unit cost by drug class**  
   - Prices were obtained from published literature (DOI: 10.1513/AnnalsATS.202008-1082RL). Prices of SABA and LAMA were digitized from Figure 2A as these prices were not reported in the text.
   - It was assumed that one inhaler is used per month per drug class, resulting in 12 inhalers annually.  
   - Annual cost per drug class (2018 Costs USD) was calculated as:  
     `Annual Cost = Inhaler Unit Price × 12`

2. **Real-world dispensing frequency based on proportion of days covered (PDC)**  
   - PDC estimates were used to determine adherence to inhaler therapies
   - Sources for PDC data include:
     - **Mannino et al. 2022** (DOI: 10.1016/j.rmed.2022.106807)
     - **Slade et al. 2021** (DOI: 10.1186/s12890-021-01612-5)
     - **Bengtson et al. 2018** (DOI: 10.1177/1753466618772750) 

#### Dispense frequency per year (adjusted via PDC)

**Mannino et al. 2022:**  
- **ICS + LAMA + LABA**: PDC = 0.66  

**Slade et al. 2021:**  
- **LAMA + LABA**: PDC = 0.44  
- **LAMA**: PDC = 0.37  

**Bengtson et al. 2018**
- **SABA**: The study reported an average of 1 fill per month 

**Adherence:** Adherence was determined to be an average of the PDC reported between the studies that assessed non-SABA inhalers= 0.49.

```{r, eval = TRUE, echo = FALSE}

library(knitr)

drug_costs <- data.frame(
  class <- c("ICS + LAMA + LABA", "LAMA", "LAMA+LABA", "SABA"),
  monthly_cost <- c(296.11, 208.10, 218.05, 32.2),
  dispenses_per_year = c(12, 12, 12, 12)
)

drug_costs$annual_cost <- round(drug_costs$monthly_cost * drug_costs$dispenses_per_year, 2)

kable(drug_costs,
      col.names = c("Drug Class", "Monthly Cost (USD)", "Dispenses/Year", "Estimated Annual Cost (USD)"),
      format.args = list(big.mark = ","),
      caption = "Estimated Annual COPD Inhaler Costs by Drug Class")

```

## Step 2: Estimating COPD-related background costs by GOLD stage

COPD-related background costs were estimated using data from Wallace et al 2019 (DOI: 10.18553/jmcp.2019.25.2.205) Table 3, specifically the row labeled "COPD-related costs, all patients". Background costs were calculated by subtracting the costs of **Inpatient Care**, **Emergency Room (ER) Visits**, and **Pharmacy** from the **Total COPD-related Medical Costs** (2016 Costs USD) as these costs are determined separately.

```{r, eval = TRUE, echo = FALSE}

copd_background_costs <- data.frame(
  `GOLD Stage` = c("GOLD I", "GOLD II", "GOLD III", "GOLD IV"),
  `Total COPD Cost` = c(5945, 6978, 10751, 18070),
  `Inpatient` = c(3853, 4449, 6277, 12139),
  `ER Visits` = c(186, 144, 193, 534),
  `Pharmacy` = c(592, 1101, 2000, 2479),
  `Background Cost` = c(1314, 1284, 2281, 2918)
)

kable(copd_background_costs,
      col.names = c("GOLD Stage", "Total COPD-related Medical Costs", "Inpatient", "ER Visits", "Pharmacy", "Background Cost (USD)"),
      format.args = list(big.mark = ","),
      caption = "COPD-related Background Costs by GOLD Stage")

```

## Step 3: Estimating exacerbation costs by severity

The exacerbation module assigns per-event direct medical costs based on severity: **Mild**, **Moderate**, **Severe**, and **Very Severe**. These cost estimates were derived from U.S.-based healthcare utilization studies.

- **Mild Exacerbation**: Defined as an increase in bronchodilator use that does not result in a healthcare encounter. It is assumed that half of the number of available doses in a SABA inhaler is used per event. Cost of SABA inhaler is obtained from the published literature (DOI: 10.1513/AnnalsATS.202008-1082RL).
- **Moderate Exacerbation**: Involves a visit to a healthcare facility (e.g., physician office or emergency department) without resulting in hospital admission.
- **Severe Exacerbation**: Defined as an inpatient hospital admission
- **Very Severe Exacerbation**: Defined as ICU admission

The following references were used:

- **Dalal et al. 2011** (DOI: 10.1016/j.rmed.2010.09.003) — 2008 Costs USD  
- **Bogart et al. 2020** (DOI: 10.37765/ajmc.2020.43157) — 2017 Costs USD

```{r, eval = TRUE, echo = FALSE}

exacerbation_costs <- data.frame(
  `Exacerbation Severity` = c("Mild", "Moderate", "Severe", "Very Severe"),
  `Definition` = c(
    "Increased SABA medication usage",
    "No hospitalization",
    "Inpatient hospitalization",
    "ICU + intubation"
  ),
  `Cost (USD)` = c(16.1, 2107, 22729, 44909),
  `Source` = c(
    "Assumption",
    "Bogart et al. 2020",
    "Bogart et al. 2020",
    "Dalal et al. 2011"
  )
)

kable(exacerbation_costs,
      col.names = c("Exacerbation Severity", "Definition", "Cost (USD)", "Source"),
      format.args = list(big.mark = ","),
      caption = "Per-Event COPD Exacerbation Costs by Severity")
```


## Step 4: Estimating costs for smoking cessation 

## Estimating the Cost of Smoking Cessation Therapy

To estimate the cost of smoking cessation therapy, the distribution of commonly used pharmacologic and behavioral therapies was obtained from the MMWR study (DOI: 10.15585/mmwr.mm7329a1). The reported usage among individuals attempting to quit included:

- **Nicotine patch**: 19.6%  
- **Nicotine gum/lozenge**: 18.4%  
- **Nicotine spray/inhaler**: 1.0%  
- **Varenicline**: 9.6%  
- **Bupropion**: 6.4%  
- **Behavioral counseling**: 7.3%  

### Reweighted to Assume 100% Uptake

To standardize the distribution across therapies, reweighting was performed using the total of all therapies (62.3%):

- **Nicotine patch**: (19.6 / 62.3) × 100 ≈ **31.5%**  
- **Nicotine gum/lozenge**: (18.4 / 62.3) × 100 ≈ **29.5%**  
- **Nicotine spray/inhaler**: (1.0 / 62.3) × 100 ≈ **1.6%**  
- **Varenicline**: (9.6 / 62.3) × 100 ≈ **15.4%**  
- **Bupropion**: (6.4 / 62.3) × 100 ≈ **10.3%**  
- **Behavioral counseling**: (7.3 / 62.3) × 100 ≈ **11.7%**  

### Estimated 3-Month Cost of Smoking Cessation Therapy

Cost estimates were based on 2025 GoodRx prices for a 3-month course of pharmacotherapy.  
For behavioral counseling, Medicare reimburses up to 8 sessions per year, which was assumed to be the number of sessions used.  
The per-session cost was based on **2015 CPT codes**:

- **99406**: \$14.37 for sessions <10 minutes  
- **99407**: \$27.67 for sessions >10 minutes  

A midpoint value of **\$21.02** per session was applied, totaling **\$168.16** for 8 sessions.

Using the reweighted proportions, the weighted average cost for 3 months of smoking cessation pharmacotherapy was calculated as:

**Average cost**: \$125.65

```{r, eval = TRUE, echo = FALSE}

smoking_cessation <- data.frame(
  Therapy = c("Nicotine Patch", "Nicotine Gum/Lozenge", "Nicotine Spray/Inhaler", 
              "Varenicline", "Bupropion", "Behavioral Counseling", "Average (weighted)"),
  `Reweighted Proportion (%)` = c(31.5, 29.5, 1.6, 15.4, 10.3, 11.7, "--"),
  `Cost (USD)` = c(71, 35, 550, 402, 25, 168.16, 125.65)
)

kable(smoking_cessation,
      col.names = c("Therapy", "Reweighted Proportion (%)", "Cost (USD)"),
      format.args = list(big.mark = ",", nsmall = 2),
      caption = "Smoking Cessation Therapy Use and Cost Estimates (3-Month Duration)")

```

## Step 5: Estimating costs for GP visits and diagnostic spirometry 

**GP visits**.  
CPT code 99214 is a standard outpatient GP visit, the midpoint between facility and non facility visit reimbursement amounts was used which equates $94.15 (2015 Costs USD)

**Spirometry**. 
CPT code 94060 was used which equates to $61.81 (2015 Costs USD)
