---
title: "RecalibratingEPIC"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RecalibratingEPIC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(epicR)
```

### Current Calibration

Assessing calibration of exacerbations with current EPIC's equations:

```{r validation}
inputs <- get_input()$values
validate_exacerbation(5e4, inputs)
```

The current input equation for exacerbations include a term for diagnosis: `r inputs$exacerbation$ln_rate_betas`

As a result, whether a person is diagnosed or not will affect the exacerbation rates, which does not have causal face validity, and is particularly undesirable for case detection studies that assess diagnosis.

If we remove the diagnosis term and go back to the previous calibration, it will look like this:

```{r validationNewEquation}
inputs$exacerbation$ln_rate_betas <- t(as.matrix(c(intercept = 1.7, female = 0, age = 0.04082 * 0.1, fev1 = -1.5, smoking_status = 0, gold1 = 0.6 , gold2 = 0.35 , gold3 = 0.08 , gold4 = -0.35 , diagnosis_effect = 0)))
validate_exacerbation(5e4, inputs)
```

### Recalibration 1: 

Let's try to improve that by lowering GOLD coefficients:

```{r validationNewEquation1}
inputs$exacerbation$ln_rate_betas <- t(as.matrix(c(intercept = 1.7, female = 0, age = 0.04082 * 0.1, fev1 = -1.5, smoking_status = 0, gold1 = 0.3 , gold2 = 0.1 , gold3 = 0.08 , gold4 = -0.35 , diagnosis_effect = 0)))
validate_exacerbation(5e4, inputs)
```

### Recalibration 2:

We can now lower GOLD2 coefficient even more and perhaps increase smoking coefficient to compensate for the loss in the diagnosed patients.

```{r validationNewEquation2}
inputs$exacerbation$ln_rate_betas <- t(as.matrix(c(intercept = 1.7, female = 0, age = 0.04082 * 0.1, fev1 = -1.5, smoking_status = 0.5, gold1 = 0.3 , gold2 = 0.05 , gold3 = 0.08 , gold4 = -0.35 , diagnosis_effect = 0)))
validate_exacerbation(5e4, inputs)
```

### Recalibration 3: 

Let's lower GOLD2 dramatically:

```{r validationNewEquation3}
inputs$exacerbation$ln_rate_betas <- t(as.matrix(c(intercept = 1.7, female = 0, age = 0.04082 * 0.1, fev1 = -1.5, smoking_status = 0.5, gold1 = 0.3 , gold2 = -0.5 , gold3 = 0.08 , gold4 = -0.35 , diagnosis_effect = 0)))
validate_exacerbation(5e4, inputs)
```

### Recalibration 4: 

Bring GOLD2 up a bit:

```{r validationNewEquation4}
inputs$exacerbation$ln_rate_betas <- t(as.matrix(c(intercept = 1.7, female = 0, age = 0.04082 * 0.1, fev1 = -1.5, smoking_status = 0.5, gold1 = 0.3 , gold2 = -0.2 , gold3 = 0.08 , gold4 = -0.35 , diagnosis_effect = 0)))
validate_exacerbation(5e4, inputs)
```

### Recalibration 5: 

GOLD2 back a bit again.

```{r validationNewEquation5}
inputs$exacerbation$ln_rate_betas <- t(as.matrix(c(intercept = 1.7, female = 0, age = 0.04082 * 0.1, fev1 = -1.5, smoking_status = 0.5, gold1 = 0.3 , gold2 = -0.3 , gold3 = 0.08 , gold4 = -0.35 , diagnosis_effect = 0)))
validate_exacerbation(5e4, inputs)
```

### Recalibration 6:

Reducing intercept and adding to smoking term.

```{r validationNewEquation5}
inputs$exacerbation$ln_rate_betas <- t(as.matrix(c(intercept = 1.4, female = 0, age = 0.04082 * 0.1, fev1 = -1.5, smoking_status = 0.7, gold1 = 0.3 , gold2 = -0.3 , gold3 = 0.08 , gold4 = -0.35 , diagnosis_effect = 0)))
validate_exacerbation(5e4, inputs)
```

### 
