---
title: "Using EPIC in R"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
vignette: >
  %\VignetteIndexEntry{UsingEPICinR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8} 
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(rmarkdown.html_vignette.check_title = FALSE)

```

## 1 Installation

If you do not have R installed on your computer, please go to [Appendix 1](#appendix1). Once you have R installed, you can install the **epicR** package as per the instructions below:

### Windows 7 or Later {-}
1. Download and Install the latest version of R from [https://cran.r-project.org/bin/windows/base/](https://cran.r-project.org/bin/windows/base/)
2. Download and Install R Studio from [https://www.rstudio.com/products/rstudio/download/](https://www.rstudio.com/products/rstudio/download/)
3. Download and Install the latest version of Rtools from [https://cran.r-project.org/bin/windows/Rtools/](https://cran.r-project.org/bin/windows/Rtools/)
4. Using either an R session in Terminal or in R Studio, install the package `remotes``:

```{r, eval=FALSE, echo=TRUE}
install.packages('remotes')
```

5. Install epicR from GitHub:

```{r, eval=FALSE, echo=TRUE}
remotes::install_github('resplab/epicR')
```

### Mac OS Sierra and Later {-}
TBD

### Ubuntu 22.04 and Later {-}

1. Install R by executing the following commands in Terminal:

```bash
  sudo apt-key adv --keyserver keyserver.ubuntu.com --recv-keys E298A3A825C0D65DFD57CBB651716619E084DAB9
```
```bash
  sudo add-apt-repository 'deb [arch=amd64,i386] https://cran.rstudio.com/bin/linux/ubuntu xenial/'
```
```bash
  sudo apt-get update
```
```bash
  sudo apt-get install r-base
```
If the installation is successful, you should be able to start R:
```bash
  sudo -i R
```

2. Download and Install R Studio from [https://www.rstudio.com/products/rstudio/download/](https://www.rstudio.com/products/rstudio/download/)
3. Install `libcurl` from Terminal: 

```bash
  sudo apt-get install libcurl4-openssl-dev libssl-dev r-base-dev
```

4. Using either an R session in Terminal or in R Studio, install the package `devtools`:

```r
install.packages ('remotes')
```
  
5. Install epicR from GitHub:

```r
remotes::install_github('resplab/epicR')
```

## 2 Running the Model

Now that you have installed `epicR`, you can load the library:

```{r, eval = FALSE, echo = TRUE}
library(epicR)
```

### Step 1: Initialize the Session {-}

The backend of this package is in C++. In other languages such as R/Python, memory allocation is taken care of for you. However, in C/C++, you need to allocate memory yourself. You also need to deallocate memory when the program is done. Because of this, we have the function `init_session()`.

`init_session()` is in the **core.R** file, and calls 3 C functions: `Cdeallocate_resources()`,
`Callocate_resources()`, and `Cinit_session()`. You don't really need to know how these work, but basically they allocate/deallocate memory.

You need to call this before you run anything:

```{r, eval = FALSE, echo = TRUE}
init_session()
```

In C++, a program that is successful returns `0`, so you should see this in the R console.

### Step 2: Set the Inputs {-}

The default input values are created in the file **input.R**. You can retrieve these values using `get_input()`. This function returns a nested list. The returned list contains `values`, `help` and `ref` which each hold further sub-lists. `help` and `ref` contain explanations regarding what each input parameter means. 

For easier notation we will refer the retrieved inputs as `input`:

```{r, eval = FALSE, echo = TRUE}
input <- get_input()
```

Below is a list of the parameters that `input$values` returns. 

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
includeCSS("html/section-3/style-table.css")
rawHTML <- paste(readLines("html/section-3/set-inputs.html"), collapse="\n")
HTML(rawHTML)
```

`input$values` returns a list of lists, so you can further explore and set specific inputs using the `$` operator. An example of how to do this is shown below:

```{r, eval = FALSE, echo = TRUE}
#changing time_horizon which can be accessed through global_parameters
input$values$global_parameters$time_horizon <- 5
#changing age0, which can be accessed through global_parameters, from the default value of 40 to 50
input$values$global_parameters$age0 <- 50
#changing p_females, which can be accessed through agent, from the default value 0.5 to 0.6.
input$values$agent$p_female <- 0.6
```


### Step 3: Run {-}

After changing the inputs you can provide them to the `run()` function. The `run()` function has two arguments: **max_n_agents** and **input**. If you created your own input in **Step 2**, you can put it in here; otherwise, you can leave the **input** argument blank. The **max_n_agents** argument is how many people you want in your model. The default is set to the maximum integer your computer will allow. You can set this to a lower number if you like (it might run faster).

```{r, eval=FALSE, echo=TRUE}
#to run the simulation with default parameters
run()
#to run the simulation with custom parameters
run(max_n_agents = 1000, input = input$values)
```

### Step 4: Results {-}

Once you have run the model simulation, there are several functions to access the results. 

The first function is the `Cget_output()` which shows some of the overall results:

```{r, eval=FALSE, echo=TRUE}
results <- Cget_output()
```

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
includeCSS("html/section-3/style-table.css")
rawHTML <- paste(readLines("html/section-3/main-results.html"), collapse="\n")
HTML(rawHTML)
```

The second function `Cget_output_ex()` returns a very large list of results:

```{r, eval=FALSE, echo=TRUE}
resultsExra <- Cget_output_ex()
```

**While interpreting the `age` column, it is important to note that EPIC only starts simulating patients at age 40. Also comorbidities are not implemented in the current version of epicR.**

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
includeCSS("html/section-3/style-table.css")
rawHTML <- paste(readLines("html/section-3/ex-results.html"), collapse="\n")
HTML(rawHTML)
```

The full snippet of the code:

```{r, eval=FALSE, echo=TRUE}
init_session()
input <- get_input()
input$values$global_parameters$time_horizon <- 5
run(input=input$values)
results <- Cget_output()
resultsExra <- Cget_output_ex()
terminate_session()
```

For some studies, having access to the entire event history of the simulated population might be beneficial. Capturing event history is possible by setting record_mode as a setting.

`Cget_all_events_matrix()` function returns the events matrix, with every event for every patient. Note that you might need a large amount of memory available, if you want to collect event history for a large number of patients.


``` {r, eval=FALSE, echo=TRUE}
settings <- get_default_settings()
# record_mode = 2 indicates recording every event that occure
settings$record_mode <- 2
#n_base_agents is the number of people at time 0.
settings$n_base_agents <- 1e4
init_session(settings = settings)
run()
results <- Cget_output()
events <- as.data.frame(Cget_all_events_matrix())
head(events)
terminate_session()
```

You can change the record mode of the simulation by accessing `settings$record_mode` as shown above. Here are what different record modes mean 


```{r, eval=TRUE, echo=FALSE}
library(htmltools)
includeCSS("html/section-3/style-mini-table.css")
rawHTML <- paste(readLines("html/section-3/record_mode.html"), collapse="\n")
HTML(rawHTML)
```

As shown in the code snippet above, you can inspect the event matrix `Cget_all_events_matrix()` by converting it into a data frame using `as.data.frame(Cget_all_events_matrix())`. This data frame consists of 33 columns, as detailed below.

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
includeCSS("html/section-3/style-table.css")
rawHTML <- paste(readLines("html/section-3/event-matrix-cols.html"), collapse="\n")
HTML(rawHTML)
```


In the events data frame, each type of event has a code corresponding to the table below:

Note: Doctor visit and Medication change are not implemented in this version if epicR

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
includeCSS("html/section-3/style-mini-table.css")
rawHTML <- paste(readLines("html/section-3/events.html"), collapse="\n")
HTML(rawHTML)
```


### Closed-cohort analysis {-}

Closed-cohort analysis can be specified by changing the appropriate input parameters:

```{r, eval=FALSE, echo=TRUE}
library(epicR)
input <- get_input(closed_cohort = 1)$values
init_session()
run(input=input)
Cget_output()
terminate_session()
```
